---
title: "rgee Review"
output:
    html_notebook:
        toc: true
        toc_float: true
editor_options:
  chunk_output_type: inline
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(magrittr)
library(devtools)
```

# `rgeeExtra` - package review

## **Reviewer:** [\@zackarno](https://github.com/zackarno)

### Review Submitted:
**`r cat(sprintf("**Last updated:** %s", Sys.Date()))`**



<br>

This report contains documents associated with the review of **rOpenSci** submitted package:

### **`rgeeExtra`: ropensci/software-review**  issue [\#NA](https://github.com/ropensci/onboarding/issues/).

<br>

## Package info

**Description:**

High-level API to process Google Earth Engine (GEE) raster (ee.Image and ee.ImageCollection) and vector data (ee.Geometry, ee.Feature, and ee.FeatureCollection). Popular Third-party GEE algorithms are re-coded from Javascript and Python to R.

**Author:** `r 
    c(person(given = "Cesar",
           family = "Aybar",
           role = c("aut", "cre"),
           email = "csaybar@gmail.com",
           comment = c(ORCID = "0000-0003-2745-9535")),
      person(given = "David",
           family = "Montero",
	         role = c("ctb"),
	         email = "dml.mont@gmail.com",
	         comment = c(ORCID = "0000-0002-9010-3286")),
      person(given = "Lesly",
           family = "Bautista",
	         role = c("ctb"),
	         email = "leslyarcelly.213@gmail.com",
	         comment = c(ORCID = "0000-0003-3523-8687")),
      person(given = "Marc",
           family = "Choisy",
	         role = c("ctb"),
	         email = "mchoisy@oucru.org",
	         comment = c(ORCID = "0000-0002-5187-6390"))
    )`

**repo url:** <https://github.com/r-earthengine/rgeeExtra>

**website url:** <https://r-earthengine.github.io/rgeeExtra/>

## Review info

**key review checks:**

- Does the code comply with **general principles in the [Mozilla reviewing guide](https://mozillascience.github.io/codeReview/review.html)**?
- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
  +  from above link: "We strongly recommend short, descriptive names in lower case"
- Are there **improvements** that could be made to the **code style?**
  + generally snake_case recommendation not followed, but I think that is done purposefully to 
  mimic GEE JavaScript syntax
  + could clean up some sections of code that use `=` rather than `<-` for assignment
- Is there **code duplication** in the package that should be reduced?
- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?
- Is the [**documentation**](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) **clear and sufficient**?

Please be respectful and kind to the authors in your reviews. The rOpenSci [code of conduct](https://ropensci.github.io/dev_guide/policies.html#code-of-conduct) is mandatory for everyone involved in our review process.

***

### session info


```{r sessionInfo}
sessionInfo()
```


```{r pkg_dir, echo = F}
pkg_dir <- "/Users/zackarno/Documents/CHD/repos/rOpenSci/reviews/rgeeExtra"
```

## Test installation

### test local `rgeeExtra` install:

```{r test-local}
install(pkg_dir, dependencies = T, build_vignettes = T)
```

```{r github-rm}
remove.packages("rgeeExtra")
```
#### **comments:**

Local installation gave this warning:

```{r}
#> The downloaded binary packages are in
#> 	/var/folders/61/cp06zhcj4y76q7rfx0qlm06c0000gn/T//RtmpKpjkIO/downloaded_packages
#> Warning: internal error -5 in R_decompress1Error in install(pkg_dir, dependencies = T, build_vignettes = T) : 
#>  lazy-load database '/Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library/pkgbuild/R/pkgbuild.rdb' is corrupt
```

It seems to have failed because:
```{r}
library(rgeeExtra)
#> Error in library(rgeeExtra) : there is no package called ‘rgeeExtra’
```

- Therefore, I clicked uninstall `{pkgbuild}` in the Rstudio package manager and repeated the install - Same warning
- Next I just tried `install.packages("pkgbuild") - after this was successful, installing rgeeExtra from local and GitHub both worked


### test install of `rgeeExtra` from GitHub with:

```{r test-github}
devtools::install_github("r-earthengine/rgeeExtra", dependencies = T, build_vignettes = T)
```


#### **comments:**

running this directly after the local install attempt yields:

```{r}
devtools::install_github("r-earthengine/rgeeExtra", dependencies = T, build_vignettes = T)

#> Warning: internal error -3 in R_decompress1Error in devtools::install_github("r-earthengine/rgeeExtra", dependencies = T,  : 
#>  lazy-load database '/Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library/pkgbuild/R/pkgbuild.rdb' is corrupt

```
I later restarted R and reinstalled to get this error:
```{r}
#> Using github PAT from envvar GITHUB_TOKEN
#> Downloading GitHub repo r-earthengine/rgeeExtra@HEAD
#> Error in utils::download.file(url, path, method = method, quiet = quiet,  : 
#>   download from 'https://api.github.com/repos/r-earthengine/rgeeExtra/tarball/HEAD' failed
```

***

## Check package integrity

### run checks on `rgeeExtra` source:

```{r check-checks}
devtools::check(pkg_dir)
```
#### **comments:**

`devtools::check(pkg_dir)`
<!-- record comments on checks here -->
0 errors ✔ | 1 warning ✖ | 1 note ✖

checking for code/documentation mismatches ... WARNING
  Variables with usage in documentation object 'ee-citation' but not in code:
    ‘ee$Image$Extra_getCitation(x)’
    ‘ee$ImageCollection$Extra_getCitation(x)’
  
  Variables with usage in documentation object 'ee-getdoi' but not in code:
    ‘ee$Image$Extra_getDOI(x)’ ‘ee$ImageCollection$Extra_getDOI(x)’
  
  Variables with usage in documentation object 'ee-getoffset' but not in code:
    ‘ee$Image$Extra_getOffsetParams(x)’
    ‘ee$ImageCollection$Extra_getOffsetParams(x)’
  
  Variables with usage in documentation object 'ee-getscaleparams' but not in code:
    ‘ee$Image$Extra_getScaleParams(x)’
    ‘ee$ImageCollection$Extra_getScaleParams(x)’
  
  Variables with usage in documentation object 'ee-getstac' but not in code:
    ‘ee$Image$Extra_getSTAC(x)’ ‘ee$ImageCollection$Extra_getSTAC(x)’
  
  Variables with usage in documentation object 'ee-pansharpen' but not in code:
    ‘ee$Image$Extra_panSharpen(x, ...)’
  
  Variables with usage in documentation object 'ee-preprocess' but not in code:
    ‘ee$Image$Extra_preprocess(x, ...)’
    ‘ee$ImageCollection$Extra_preprocess(x, ...)’
  
  Variables with usage in documentation object 'ee-scaleandoffset' but not in code:
    ‘ee$Image$Extra_scaleAndOffset(x)’
    ‘ee$ImageCollection$Extra_scaleAndOffset(x)’
  
  Variables with usage in documentation object 'ee-tasseledcap' but not in code:
    ‘ee$Image$Extra_tasseledCap(x)’
  
  Variables with usage in documentation object 'ee_ImageCollection_closest' but not in code:
    ‘ee$ImageCollection$Extra_closest(x, ...)’
  
  Variables with usage in documentation object 'ee_Image_maskClouds' but not in code:
    ‘ee$Image$Extra_maskClouds(x, ...)’
  
  Variables with usage in documentation object 'ee_Image_matchHistogram' but not in code:
    ‘ee$Image$Extra_matchHistogram(x, ...)’
  
  Variables with usage in documentation object 'ee_Image_spectralIndex' but not in code:
    ‘ee$Image$Extra_spectralIndex(x, ...)’
  
  Variables with usage in documentation object 'extremeValues' but not in code:
    ‘ee$Image$Extra_maxValue(x, ...)’ ‘ee$Image$Extra_minValue(x, ...)’
  
  Variables with usage in documentation object 'names(x) <- value' but not in code:
    ‘names(x) <- value’
  
  Variables with usage in documentation object 'names.ee.image.Image' but not in code:
    ‘names(x)’


checking installed package size ... NOTE
    installed size is 15.2Mb
    sub-directories of 1Mb or more:
      doc       12.4Mb
      ee_extra   1.6Mb

***

### run tests on `rgeeExtra` source:

```{r check-tests}
devtools::test(pkg_dir)
```
#### **comments:**

══ Results ═══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════
Duration: 273.4 s

[ FAIL 0 | WARN 0 | SKIP 0 | PASS 135 ]

***


### check `rgeeExtra` for goodpractice:

```{r test-goodpractice}
goodpractice::gp(pkg_dir)
```
#### **comments:**
tests
── GP rgeeExtra ──────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────

It is good practice to

  ✖ write unit tests for all functions, and all package code in general. 90% of code lines are covered by test
    cases.

    R/ee_gif.R:230:NA
    R/ee_gif.R:231:NA
    R/ee_gif.R:232:NA
    R/ee_gif.R:233:NA
    R/ee_gif.R:234:NA
    ... and 46 more lines




## Check package metadata files

### inspect

- #### [README](https://github.com/r-earthengine/rgeeExtra)
- #### [DESCRIPTION](https://github.com/r-earthengine/rgeeExtra/blob/master/DESCRIPTION)
- #### [NAMESPACE](https://github.com/r-earthengine/rgeeExtra/blob/master/NAMESPACE)

### spell check

```{r spell-check}
devtools::spell_check(pkg_dir)
```


#### **comments:**

`devtools::spell_check()` results that seem worth fixing

Aditional            Aditional.Rmd:2
IGARSS’            ee-tasseledcap.Rd:50
coefficeints         ee-tasseledcap.Rd:43

***

## Check documentation

online documentation: **<https://r-earthengine.github.io/rgeeExtra/>**

* Is the [documentation](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) clear and sufficient?

### test `rgeeExtra` function help files:

```{r test-help}
help(package = "rgeeExtra")
```

#### **comments:**
- working now
`help(package = "rgeeExtra")`return "URL '/help/library/rgeeExtra/html/00Index.html' not found" in Rstudio Help viewer



### test `rgeeExtra` vignettes:

```{r test-vignettes}
vignette(package = "rgeeExtra")
```

#### **comments:**

Loaded well

***

## Test functionality:

- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?

```{r free-style}
library("rgeeExtra")
```

```{r parse-functions}
exports <-ls("package:rgeeExtra")
exports
```

## Overall Comments

- is there a way to make it easier for users to access the function help panel (i.e ?func_name). For example how do users find out more information about the `Extra_*` functions from within Rstudio. In `{rgee}` we can access the documentation with `ee_help()`, but that doesn't work here. After quite a bit of searching I eventually realized there are unexported functions (i.e ee_Image_maskClouds,ee_Image_matchHistogram , etc ) with the help documentation, but this doesn't seem like a super user-friendly place to keep it?
- enhancement idea: add a `list_spectral_indices()` type function (like eemont.listIndices())- would be cool if the object returned would provide the formula & reference for the index as well
- mention [ee_extra](https://github.com/r-earthengine/ee_extra) package somewhere in readme?
- a fair amount of functions freeze R session - any checks warnings that could be put in place?

```{r}
library(rgee)
library(rgeeExtra)
library(tidyverse)
library(sf)
ee_Initialize()
extra_Initialize()
url_s2 <- "COPERNICUS/S2"
ic <- ee$ImageCollection(url_s2)

```

## `ee_ImageCollection_closest`/ `Extra_closest()` (3 issues/1 feature idea)
- **1. date issue** doesn't accept an r Date (as_date(x)) as specified in ?help doc (date issue)
- **2. return value issue** - Return ImageCollection, contrary to return value specified in ?help doc. If area just contains 1 closest image then perhaps it should return image? Otherwise image collection? (return)
- **3. error warning/message issue** See error message in code below
- be nice if there was an `n` argument to return more than 1 image


### Date issue
`?ee_ImageCollection_closest` states: "dateee$Date or R date object. Date of interest for searching the closest image." But doesn't seem to work with R date. Also is that a typo in the documentation **dateee$Date**?

```{r}
date_chr <- "2023-09-10"
date_dt<- as_date(date_chr)

ic_closest_from_dt_obj <-  ic$Extra_closest(date_dt)
#> Error in py_call_impl(callable, call_args$unnamed, call_args$named) : 
#>   ee.ee_exception.EEException: Invalid argument specified for ee.Date(): 2023-09-10
#> Run `reticulate::py_last_error()` for details.

```

what happens whe **NO** AOI is defined?
```{r}
ic_closest <- ic$Extra_closest(date_chr)

# this is running for a loooong time and always crahses
ic_closest_dates <- ee_get_date_ic(ic_closest)
ic_closest_dates
```

When we define an AOI
```{r}

# derna Libya
ee_aoi <- ee$Geometry$Point( c(22.62697 ,32.75681))

ic_bounded <- ic$filterBounds(ee_aoi)

```



```{r}
# okay lets try closest date

 # did this 2x in a row and got the
ic_bounded_closest <- ic_bounded$Extra_closest(date_chr)
ic_bounded_closest %>% class() # interesting that it is an IC... Shouldn't it be an image?

ee_get_date_ic(ic_bounded_closest) 
```


- Eventually I did the above and it worked. Is that something with my connection? Was it my connection on the `Extra_closest` call or was it on `ee_get_date_ic()`... either way would be good to have a better error message.

### Return Value Issue

```{r}
ic_bounded_closest_5d <- ic_bounded$Extra_closest(date_chr, 5,"day")
ee_get_date_ic(ic_bounded_closest_5d)
```


`?ee_ImageCollection_closest shows:` return "Value: An ee$Image closest to the specified date." - but it doesn't return an image.



## rgeeExtra::ee_get()
- wonder if this name is a bit ambiguous considering all the `get_*` functions there are in `rgee`,`GEE` and `rgeeExtra`... a couple ideas:
- somethings with filter (filterIndex, filter_index)?
- just ic_idx?
- something with Subset ... subset_index?



# Extra_spectralIndex

```{r}
# Load first COPERNICUS/S2_SR image, compute NDVI and SAVI
s2_indices <- ee$ImageCollection("COPERNICUS/S2_SR")[[1]] %>%
  ee$Image$Extra_preprocess() %>%
  ee$Image$Extra_spectralIndex(c("NDVI", "SAVI"))

# Set visualization parameters
visParams <- list(
  min = 0, 
  max = 0.2, 
  palette = c("brown", "yellow", "green")
)

# Display NDVI on map with custom visuals
Map$centerObject(s2_indices, zoom = 10)
Map$addLayer(
  eeObject = s2_indices[["NDVI"]], 
  visParams = visParams,
  name = "NDVI"
)

```

## Simple Band Math

```{r}
# Compute squared NDVI from Landsat 8 image
img <- ee$Image("LANDSAT/LC08/C02/T1_L2/LC08_007067_20140822")
ndvi <- ((img[["SR_B5"]] - img[["SR_B4"]]) / (img[["SR_B5"]] + img[["SR_B4"]])) ** 2
names(ndvi) <- "NDVI"

# Visualize squared NDVI on map
Map$centerObject(ndvi)
Map$addLayer(
  eeObject = ndvi, 
  visParams = list(
    min = 0, 
    max = 0.2, 
    palette = c("brown", "yellow", "green")
    ),
  name = "Squared NDVI"
  )
```


## Cloud Masking
```{r exp-chunk}
img <- ee$ImageCollection("COPERNICUS/S2_SR") %>%
  ee$ImageCollection$first() %>%
  ee$Image$Extra_maskClouds(prob = 75,buffer = 300,cdi = -0.5)

names(img)
```



# Aditional Vignette
None of the code in this [aditional vignette](https://github.com/r-earthengine/rgeeExtra/blob/master/vignettes/Aditional.Rmd#L33) would run on my computer. I think it's because all the examples are missing there `Extra_` prefix (also: Aditional -> Additional) 


Below I've added them so that I could get passed the errors and test them out
## getOffsetParams()

The very first time I ran this I got this error;
```{r}
ee$ImageCollection("NASA/GPM_L3/IMERG_V06") %>%
  ee$ImageCollection$first() %>%
  ee$ImageCollection$Extra_getOffsetParams()
```



## getScaleParams()
```{r}
ee$ImageCollection("NASA/GPM_L3/IMERG_V06") %>%
  ee$ImageCollection$first() %>%
  ee$Image$Extra_getScaleParams()

?rgeeExtra:::Extra_getScaleParams

```

## Spectral Indices

```{r}


s2_indices <- ee$ImageCollection("COPERNICUS/S2_SR") %>%
  ee$ImageCollection$first() %>%
  ee$Image$Extra_preprocess()%>%
  ee$Image$Extra_spectralIndex(c("NDVI", "SAVI"))

names(s2_indices)
```

```{r}
img <- ee$ImageCollection("COPERNICUS/S2_SR") %>%
  ee$ImageCollection$first() %>%
  ee$Image$Extra_maskClouds(prob = 75,buffer = 300,cdi = -0.5)

names(img)
```

The `getStack()` example does not work w/ `Extra_` added. I don't think the getInfo() call belongs

Original:

```{r exp-chunk}
ee$ImageCollection("NASA/GPM_L3/IMERG_V06") %>%
  ee$ImageCollection$first() %>%
  ee$Image$getSTAC() %>%
  ee$Image$getInfo()
#> Error in py_get_attr_impl(x, name, silent) : 
#>   AttributeError: type object 'Image' has no attribute 'getSTAC'
```

add getInfo()
```{r}
ee$ImageCollection("NASA/GPM_L3/IMERG_V06") %>%
  ee$ImageCollection$first() %>%
  ee$Image$Extra_getSTAC() %>%
  ee$Image$getInfo()
#> Error: unexpected symbol in "add getInfo"  
```

works w/ `Extra_` added and no getInfo() call.
```{r}
ee$ImageCollection("NASA/GPM_L3/IMERG_V06") %>%
  ee$ImageCollection$first() %>%
  ee$Image$Extra_getSTAC()
```


```{r}
ee$ImageCollection("NASA/GPM_L3/IMERG_V06") %>%
  ee$ImageCollection$first() %>%
  ee$Image$Extra_getDOI()
```


`|` operator - somehow i have `{rgee}` and `{rgeeExtras}` installed, but not leaflet.extras2

I guess the required packages should be loaded at top of script? or put a note in the vignette?
```{r}
img <- ee$Image("LANDSAT/LC08/C01/T1_TOA/LC08_047027_20160819")
img_sharp <- ee$Image$Extra_panSharpen(img, method="HPFA", qa=c("MSE", "RMSE"), maxPixels=1e13)

Map$centerObject(img)
# computer freezes 
Map$addLayer(img_sharp, list(bands=c("B4", "B3", "B2"))) |
Map$addLayer(img, list(bands=c("B4", "B3", "B2")))



#> Error in ee_check_packages("| operator", c("leaflet", "leaflet.extras2")) : 
#>  | operator required the package: leaflet.extras2. Please install it first.
```

```{r}
rgeeExtra:::load_ee_Extra
savi <- ee$Image("COPERNICUS/S2_SR/20190310T105851_20190310T110327_T30TVK") %>% 
  ee$Image$preprocess()%>%
  ee$Image$spectralIndex("SAVI")

geoviz_1 <- list(palette = c('001fff', '00ffff', 'fbff00', 'ff0000'), min=0, max=0.3)
savi[savi > 0.3] <- 0

Map$addLayer(savi, geoviz, "SAVI") | Map$addLayer(log1p(savi)*2, geoviz, "log_SAVI")
```

#### **comments:**

<!-- record comments on package experimentation here -->

***

## Inspect code:

- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
    * good [function & variable naming?](https://ropensci.github.io/dev_guide/building.html#function-and-argument-naming)
    * good [dependency management](https://ropensci.github.io/dev_guide/building.html#package-dependencies)?
- Are there **improvements** that could be made to the [**code style?**](https://ropensci.github.io/dev_guide/building.html#code-style)
- Is there **code duplication** in the package that should be reduced?

```{r inspect-code}
pkgreviewr::pkgreview_print_source("rgeeExtra")
```
**\* might not be suitable for large packages with many exported functions**


<br>
<br>

#### **comments:**

<!-- record comments on package source code here -->


## Review test suite:

See guidance on [testing](https://ropensci.github.io/dev_guide/building.html#testing) for further details.

### test coverage

```{r pkg_coverage}
covr::package_coverage(pkg_dir)

```

### inspect [tests](https://github.com/r-earthengine/rgeeExtra/blob/master/tests/testthat)


#### **comments:**

<!-- record comments on testing suite here -->


***
