---
output:
    html_notebook:
        toc: true
        toc_float: true
editor_options:
  chunk_output_type: inline
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(magrittr)
library(devtools)
```

# `rgeeExtra` - package review

## **Reviewer:** [\@zackarno](https://github.com/zackarno)

### Review Submitted:
**`r cat(sprintf("**Last updated:** %s", Sys.Date()))`**

***

<br>

This report contains documents associated with the review of **rOpenSci** submitted package:

### **`rgeeExtra`: ropensci/software-review**  issue [\#NA](https://github.com/ropensci/onboarding/issues/).

<br>

## Package info

**Description:**

High-level API to process Google Earth Engine (GEE) raster (ee.Image and ee.ImageCollection) and vector data (ee.Geometry, ee.Feature, and ee.FeatureCollection). Popular Third-party GEE algorithms are re-coded from Javascript and Python to R.

**Author:** `r 
    c(person(given = "Cesar",
           family = "Aybar",
           role = c("aut", "cre"),
           email = "csaybar@gmail.com",
           comment = c(ORCID = "0000-0003-2745-9535")),
      person(given = "David",
           family = "Montero",
	         role = c("ctb"),
	         email = "dml.mont@gmail.com",
	         comment = c(ORCID = "0000-0002-9010-3286")),
      person(given = "Lesly",
           family = "Bautista",
	         role = c("ctb"),
	         email = "leslyarcelly.213@gmail.com",
	         comment = c(ORCID = "0000-0003-3523-8687")),
      person(given = "Marc",
           family = "Choisy",
	         role = c("ctb"),
	         email = "mchoisy@oucru.org",
	         comment = c(ORCID = "0000-0002-5187-6390"))
    )`

**repo url:** <https://github.com/r-earthengine/rgeeExtra>

**website url:** <https://r-earthengine.github.io/rgeeExtra/>

## Review info


#### See [reviewer guidelines](https://ropensci.github.io/dev_guide/reviewerguide.html) for further information on the rOpenSci review process.

**key review checks:**

- Does the code comply with **general principles in the [Mozilla reviewing guide](https://mozillascience.github.io/codeReview/review.html)**?
- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
- Are there **improvements** that could be made to the **code style?**
- Is there **code duplication** in the package that should be reduced?
- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?
- Is the [**documentation**](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) **clear and sufficient**?

Please be respectful and kind to the authors in your reviews. The rOpenSci [code of conduct](https://ropensci.github.io/dev_guide/policies.html#code-of-conduct) is mandatory for everyone involved in our review process.

***

### session info


```{r sessionInfo}
sessionInfo()
```


```{r pkg_dir, echo = F}
pkg_dir <- "/Users/zackarno/Documents/CHD/repos/rOpenSci/reviews/rgeeExtra"
```

## Test installation

### test local `rgeeExtra` install:

```{r test-local}
install(pkg_dir, dependencies = T, build_vignettes = T)
```

```{r github-rm}
remove.packages("rgeeExtra")
```
#### **comments:**

Local installation gave this warning:

```{r}
#> The downloaded binary packages are in
#> 	/var/folders/61/cp06zhcj4y76q7rfx0qlm06c0000gn/T//RtmpKpjkIO/downloaded_packages
#> Warning: internal error -5 in R_decompress1Error in install(pkg_dir, dependencies = T, build_vignettes = T) : 
#>  lazy-load database '/Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library/pkgbuild/R/pkgbuild.rdb' is corrupt
```

It seems to have failed because:
```{r}
library(rgeeExtra)
#> Error in library(rgeeExtra) : there is no package called ‘rgeeExtra’
```

- Therefore, I clicked uninstall `{pkgbuild}` in the Rstudio package manager and repeated the install - Same warning
- Next I just tried `install.packages("pkgbuild") - after this was successful, installing rgeeExtra from local and GitHub both worked


### test install of `rgeeExtra` from GitHub with:

```{r test-github}
devtools::install_github("r-earthengine/rgeeExtra", dependencies = T, build_vignettes = T)
```


#### **comments:**

running this directly after the local install attempt yields:

```{r}
devtools::install_github("r-earthengine/rgeeExtra", dependencies = T, build_vignettes = T)
#> Warning: internal error -3 in R_decompress1Error in devtools::install_github("r-earthengine/rgeeExtra", dependencies = T,  : 
#>  lazy-load database '/Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library/pkgbuild/R/pkgbuild.rdb' is corrupt
```

***



## Check package integrity

### run checks on `rgeeExtra` source:

```{r check-checks}
devtools::check(pkg_dir)
```
#### **comments:**

`devtools::check(pkg_dir)`
<!-- record comments on checks here -->
0 errors ✔ | 1 warning ✖ | 1 note ✖

checking for code/documentation mismatches ... WARNING
  Variables with usage in documentation object 'ee-citation' but not in code:
    ‘ee$Image$Extra_getCitation(x)’
    ‘ee$ImageCollection$Extra_getCitation(x)’
  
  Variables with usage in documentation object 'ee-getdoi' but not in code:
    ‘ee$Image$Extra_getDOI(x)’ ‘ee$ImageCollection$Extra_getDOI(x)’
  
  Variables with usage in documentation object 'ee-getoffset' but not in code:
    ‘ee$Image$Extra_getOffsetParams(x)’
    ‘ee$ImageCollection$Extra_getOffsetParams(x)’
  
  Variables with usage in documentation object 'ee-getscaleparams' but not in code:
    ‘ee$Image$Extra_getScaleParams(x)’
    ‘ee$ImageCollection$Extra_getScaleParams(x)’
  
  Variables with usage in documentation object 'ee-getstac' but not in code:
    ‘ee$Image$Extra_getSTAC(x)’ ‘ee$ImageCollection$Extra_getSTAC(x)’
  
  Variables with usage in documentation object 'ee-pansharpen' but not in code:
    ‘ee$Image$Extra_panSharpen(x, ...)’
  
  Variables with usage in documentation object 'ee-preprocess' but not in code:
    ‘ee$Image$Extra_preprocess(x, ...)’
    ‘ee$ImageCollection$Extra_preprocess(x, ...)’
  
  Variables with usage in documentation object 'ee-scaleandoffset' but not in code:
    ‘ee$Image$Extra_scaleAndOffset(x)’
    ‘ee$ImageCollection$Extra_scaleAndOffset(x)’
  
  Variables with usage in documentation object 'ee-tasseledcap' but not in code:
    ‘ee$Image$Extra_tasseledCap(x)’
  
  Variables with usage in documentation object 'ee_ImageCollection_closest' but not in code:
    ‘ee$ImageCollection$Extra_closest(x, ...)’
  
  Variables with usage in documentation object 'ee_Image_maskClouds' but not in code:
    ‘ee$Image$Extra_maskClouds(x, ...)’
  
  Variables with usage in documentation object 'ee_Image_matchHistogram' but not in code:
    ‘ee$Image$Extra_matchHistogram(x, ...)’
  
  Variables with usage in documentation object 'ee_Image_spectralIndex' but not in code:
    ‘ee$Image$Extra_spectralIndex(x, ...)’
  
  Variables with usage in documentation object 'extremeValues' but not in code:
    ‘ee$Image$Extra_maxValue(x, ...)’ ‘ee$Image$Extra_minValue(x, ...)’
  
  Variables with usage in documentation object 'names(x) <- value' but not in code:
    ‘names(x) <- value’
  
  Variables with usage in documentation object 'names.ee.image.Image' but not in code:
    ‘names(x)’


checking installed package size ... NOTE
    installed size is 15.2Mb
    sub-directories of 1Mb or more:
      doc       12.4Mb
      ee_extra   1.6Mb

***

### run tests on `rgeeExtra` source:

```{r check-tests}
devtools::test(pkg_dir)
```
#### **comments:**

══ Results ═══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════
Duration: 273.4 s

[ FAIL 0 | WARN 0 | SKIP 0 | PASS 135 ]

***


### check `rgeeExtra` for goodpractice:

```{r test-goodpractice}
goodpractice::gp(pkg_dir)
```
#### **comments:**

<!-- record comments on goodpractice here -->

***

## Check package metadata files

### inspect

- #### [README](https://github.com/r-earthengine/rgeeExtra)
- #### [DESCRIPTION](https://github.com/r-earthengine/rgeeExtra/blob/master/DESCRIPTION)
- #### [NAMESPACE](https://github.com/r-earthengine/rgeeExtra/blob/master/NAMESPACE)

### spell check

```{r spell-check}
devtools::spell_check(pkg_dir)
```

#### **comments:**

<!-- record comments on metadata files here -->

***

## Check documentation

online documentation: **<https://r-earthengine.github.io/rgeeExtra/>**

* Is the [documentation](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) clear and sufficient?

### test `rgeeExtra` function help files:

```{r test-help}
help(package = "rgeeExtra")
```

#### **comments:**

<!-- record comments on help files here -->

***

### test `rgeeExtra` vignettes:

```{r test-vignettes}
vignette(package = "rgeeExtra")
```

#### **comments:**

<!-- record comments on vignettes here -->

***

## Test functionality:

- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?

```{r free-style}
library("rgeeExtra")
```

```{r parse-functions}
exports <-ls("package:rgeeExtra")
exports
```

<!-- experiment with package functions -->

```{r exp-chunk}


```

#### **comments:**

<!-- record comments on package experimentation here -->

***

## Inspect code:

- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
    * good [function & variable naming?](https://ropensci.github.io/dev_guide/building.html#function-and-argument-naming)
    * good [dependency management](https://ropensci.github.io/dev_guide/building.html#package-dependencies)?
- Are there **improvements** that could be made to the [**code style?**](https://ropensci.github.io/dev_guide/building.html#code-style)
- Is there **code duplication** in the package that should be reduced?

```{r inspect-code}
pkgreviewr::pkgreview_print_source("rgeeExtra")
```
**\* might not be suitable for large packages with many exported functions**


<br>
<br>

#### **comments:**

<!-- record comments on package source code here -->


## Review test suite:

See guidance on [testing](https://ropensci.github.io/dev_guide/building.html#testing) for further details.

### test coverage

```{r pkg_coverage}
covr::package_coverage(pkg_dir)

```

### inspect [tests](https://github.com/r-earthengine/rgeeExtra/blob/master/tests/testthat)


#### **comments:**

<!-- record comments on testing suite here -->


***
