---
title: "review_rmd"
output: html_document
date: "2024-01-04"
---
---
title: "review"
output: 
  rmarkdown::md_document:
    pandoc_args: [
      "--wrap=none"
    ]
---
```{r, echo=F, eval=T}
knitr::opts_chunk$set(echo = T,include = T, eval= F)
```


```{r ,echo=F}
library(rgee)
library(rgeeExtra)
library(tidyverse)
ee_Initialize()
extra_Initialize()
```

Thanks for the fantastic package and the chance to review! I look forward to incorporating it into my work 

I tried to limit the scope of my review to just `{rgeeExtra}` and attempted to avoid functionality of `{rgee}` and also not dig into the `EE_extra`/`EE-Mont` python packages which it wraps. 

There is so much cool stuff here that I feel like I could keep digging for a while longer, but will have to stop myself here. Below is my review:


## Package Review

*Please check off boxes as applicable, and elaborate in comments below.  Your review is not limited to these topics, as described in the reviewer guide*

- **Briefly describe any working relationship you have (had) with the package authors.**
- [x] As the reviewer I confirm that there are no [conflicts of interest](https://devguide.ropensci.org/policies.html#coi) for me to review this work (if you are unsure whether you are in conflict, please speak to your editor _before_ starting your review).

## Documentation

The package includes all the following forms of documentation:

- [x]  **A statement of need:** clearly stating problems the software is designed to solve and its target audience in README
- [x]  **Installation instructions:** for the development version of package and any non-standard dependencies in README
  + all good, but I added something as a nice-to- have below!
- [ ]  **Vignette(s):** demonstrating major functionality that runs successfully locally
  +  see below issue/bugs found in vignettes that causes them not to run properly
- [ ] **Function Documentation:** for all exported functions
  +  see `Documentation - Functions` section
- [ ] **Examples:** (that run successfully locally) for all exported functions
  +  see `Documentation - Examples` section
- [x] **Community guidelines:** including contribution guidelines in the README or CONTRIBUTING, and DESCRIPTION with `URL`, `BugReports` and `Maintainer` (which may be autogenerated via `Authors@R`).

### Documentation - Installation instructions
 
- When I am using a more typical python workflow, I feel like I have a clear view of the venv and anything im adding to it. In this setup, it seems less visible. I believe that is largely intended and makes sense for alot of R userse. Nonetheless it would be nice if there was more documentation on this for people who are more comfortable with python. For example, I would personally find it useful to see an example of how to `pip install` all the necessary dependencies for both `{rgee}` and `{rgeeExtra}` into a pre-existing/or new `venv`.

### Documentation - Examples 

- No documentation on exported mean method https://github.com/r-earthengine/rgeeExtra/blob/7ea67d064e56e9429971d64440d8424df68f6249/R/ops-math.R#L258-L264
- example provided for `Summary.ee.image.Image` returns an image with 0 pixels so I am not reall ysure what exactly it is supposed to be doing?
- Example on exported `names.ee.feature.Feature` method doesnt show function being used: https://github.com/r-earthengine/rgeeExtra/blob/7ea67d064e56e9429971d64440d8424df68f6249/R/ops-subsetting-fc.R#L57-L69
- Missing roxygen examples for `[` `[[` examples in `ops-subsetting-image.R`. It's not entirely clear what these are meant to do on an image so I think examples help
- ops-subsetting-imagecollection the 2 top funcs have the same roxygen description but doing differnt things
- `ops-subsetting.R` top example should have a `library(rgeeExtra)` call right?

### Documentation - Functions

- `ee$Image$Extra_preprocess()` - use, examples, and documentation a bit opaque.unclear what the `...` arguments should be as it references `ee$Image$cloudmask()` which I do not believe is a function (maybe you meant `Extra_maskClouds`?). Since there are different ways to cloud mask different images perhaps the documentation can give some information about the defaults and which satellites they are optimal for.
- `Extra_matchHistogram()`-  Would be useful to add `Map$addLayer()|Map$addLayer()` to roxygen example on function
- `ops-math` - roxygen in below snippet says multiplication, but shows addition 
https://github.com/r-earthengine/rgeeExtra/blob/7ea67d064e56e9429971d64440d8424df68f6249/R/ops-math.R#L40-L44
- `ee_utils_gif_save()` roxygen says `@return No return value, called to write a GIF file.` but it does return a logical after writing the gif (i'm not sure why this is necessary)
- `Extra_matchHistogram()` could be nice to provide a link and or some information on matching method

### Documentation - Vignettes

- [aditional vignette](https://github.com/r-earthengine/rgeeExtra/blob/master/vignettes/Aditional.Rmd#L33)
  + Most examples are missing the `Extra_` prefix on the functions so they result in error
  + spelling: "Aditional" -> "Additional"  :-)
- I would check other Vignettes for this same issue.

## Functionality

- [x] **Installation:** Installation succeeds as documented.
- [ ] **Functionality:** Any functional claims of the software have been confirmed.
  + See `Functionality - Functionality` section below
- [x] **Performance:** Any performance claims of the software have been confirmed.
- [x] **Automated tests:** Unit tests cover essential functions of the package and a reasonable range of inputs and conditions. All tests pass on the local machine.
- [x] **Packaging guidelines**: The package conforms to the rOpenSci packaging guidelines.

Estimated hours spent reviewing: 20

- [x] Should the author(s) deem it appropriate, I agree to be acknowledged as a package reviewer ("rev" role) in the package DESCRIPTION file.


## Functionality/ User Experience

### Functionality -Installation

- Worked fine for me. However in the past I have seen users struggle (including myself), but I believe this has gotten easier.

### Functionality - Functionality

- **Exra_Closest**- This is a very handy function! but some issues/questions
  + it returns and `ImageCollection` not an `Image` as written in the roxygen although I agree, in that, based on how it is currently constructed it would be better to return an Image (https://github.com/r-earthengine/rgeeExtra/blob/7ea67d064e56e9429971d64440d8424df68f6249/R/ee_ImageCollection.R#L18)
  + Can't take R formatted dates as written in roxygen (but should)

```{r}
library(rgee)
library(rgeeExtra)
library(tidyverse)
ee_Initialize()
extra_Initialize()

url_s2 <- "COPERNICUS/S2"
ic <- ee$ImageCollection(url_s2)

date_chr <- "2023-09-10"
date_dt<- as_date(date_chr)
ic_closest_from_dt_obj <-  ic$Extra_closest(date_dt)
#> Error in py_call_impl(callable, call_args$unnamed, call_args$named) : 
#>   ee.ee_exception.EEException: Invalid argument specified for ee.Date(): 2023-09-10
#> Run `reticulate::py_last_error()` for details.
```
  + Could not find any documentation on what is supposed to happen on geographically large ImageCollection multiple images on same date. Theoretically as a user I might think that it would give me an ImageCollection containing all the images closes to that date, but I don't think that is what you were going for. Nonetheless I think we need some instructions in the roxygen and vignettes. Also this below cases is an example where `Extra_closest()` silently fails and then crashes/aborts the session on the next call that tries to do something with the resulting object. This is a case where some assertions would improve user experience.
  
  
```{r}
ic_closest <-  ic$Extra_closest(date_chr) # works
class(ic_closest)
#> [1] "ee.imagecollection.ImageCollection" "ee.collection.Collection"           "ee.element.Element"                
#>[4] "ee.computedobject.ComputedObject"   "ee.encodable.Encodable"             "python.builtin.object"  

ee_get_date_ic(ic_closest)
```
  
  + **enhancemement idea** - would be cool if user could also supply an image instead of a date. The function then could find the closest image in the image collection to that date Like they do in the `find_closest()` function in this [example](https://developers.google.com/earth-engine/tutorials/community/histogram-matching). 
- `Extra_minValue/Extra_maxValue()` - I think this function is only designed for `ee.Image` therefore some class assertions would be nice to avoid slow "ee.ee_exception.EEException: User memory limit exceeded" message if someone runs on anything else?

- `Extra_matchHistogram()` - probably a user error, but I had a hard time trying to reproduce the [Histogram Matching Example](https://developers.google.com/earth-engine/tutorials/community/histogram-matching) from google. 
```{r}
library(rgee)
library(rgeeExtra)
library(tidyverse)
ee_Initialize()
extra_Initialize()

geometry = ee$Geometry$Polygon(
  list(
    c(-155.97117211519446, 20.09006980142336),
    c(-155.97117211519446, 19.7821681268256),
    c(-155.73256280122962, 19.7821681268256),
    c(-155.73256280122962, 20.09006980142336)
  )
)
skysat = ee$Image('SKYSAT/GEN-A/PUBLIC/ORTHO/RGB/s01_20161020T214047Z')$clip(geometry)
landsat_ic = ee$ImageCollection('LANDSAT/LC08/C01/T1_SR')$
  filterBounds(geometry) %>% 
  ee$ImageCollection$Extra_preprocess()

date_skysat <- ee_get_date_img(skysat)$time_start # if you run Extra_closest w/ this it aborts session
date_skysat_chr<- as.character(as_date(date_skysat))

landsat_ic_closest <- ee$ImageCollection$Extra_closest(landsat_ic ,date_skysat_chr,32,"day")

# would be nice if Extra_closest returns img in the above
landsat_img_closest <- landsat_ic_closest[[1]] # but alas we have to convert it

landsat_img_rgb <- landsat_img_closest$select(
  opt_selectors = c( "B4","B3","B2"),
  opt_names = c("R","G","B")
)

bands <- list("R"="R","G"="G","B"="B")
matched <- ee$Image$Extra_matchHistogram(skysat,landsat_img_rgb,  bands)

vis_params_refl = list('min'=0, 'max'= 0.25)
vis_params_dn = list('min'= 0, 'max'= 255)
aoi_center_pt <- ee$Geometry$Point(c(-155.79584,19.9986))
Map$centerObject(aoi_center_pt,13)

m1 <- Map$addLayer(landsat_img_rgb, visParams = vis_params_refl,"landsat ref")
m2 <- Map$addLayer(skysat,visParams= vis_params_dn, 'SkySat source')
m1|m2 # looks good
m3 <- Map$addLayer(matched,visParams= vis_params_refl, 'SkySat matched')
m4 <- Map$addLayer(matched,visParams= vis_params_dn, 'SkySat matched')
m1|m3 # matches is just black 
m1|m4 # matches is just black 
```

For a user like me who wrote the code below what are some tools we have from `{rgee}` and `{rgeeExtra}` to help troublshoot and figure out what is going on? This is all I can think of right now but doesn't get me much further?
```{r}
rgeeExtra::ee_minValue(landsat_img_rgb) 
rgeeExtra::ee_maxValue(landsat_img_rgb) 
rgeeExtra::ee_minValue(skysat)
rgeeExtra::ee_maxValue(skysat) 


rgeeExtra::ee_minValue(matched) # all 0s
rgeeExtra::ee_maxValue(matched) # all 0s
```



### Review Comments
## General Comments

- One thing I fund challenging about this package is that its not straight forward to get help with 
"?function_name" syntax. In `{rgee}` you are able to do it through `ee_help()` which is not as straight forwrard as ?function_name, but still better than nothing. Would there be a way to implement either ee_help() or ?function_name help here?

- It would be nice if the `How does it work` section of the readme explained a bit more about it's relation to `ee_extra`/`eemont`/`spectralIndex` python packages.
  + For most of this review I had thought you were just porting eemont functions over to rgeeExtra with reticulate, but at the end just realized that ee_extra (python) in the inst folder. Is that coming from: [ee_extra](https://github.com/r-earthengine/ee_extra) package somewhere in readme?

- There are some really nice defensive assertions you put in place on certain functions particularly the subsetting funcs. I think these really improve user experience and should be added more throughout the package. There are many cases where if the user inputs the wrong type of value the function will either a. eventually timeout, b. completely freeze/abort the session, c. give no warning and you only find out later. I've flagged a few in which some simple class assertions should help the user avoid this experience.

- `names()` - I'm sure you have thought more about this than I have so there could be a reason for not doing this, but my opinon now is it would be if `names()` was consistent across ImageCollection and Images. Currently names returns band names for Images and property names for ImageCollections. Why not just have `property_names()` and `names()` (or something like that) that works consistently since both Images an ImageCollections have both.

- would be nice to have a `list_spectral_indices()` type function (like eemont.listIndices())- would be cool if the object returned would provide the formula & reference for the index as well

- `Extra_closest()` - really cool function some comments (in addition to what I believe to be a small bug - see `Functionality-Functionality` section)
  + I think it would be very nice to be able to also provide an image rather than a date as the first argument. This would allow the user to find the closest temporally closest image to the in the ic.
  +. what happens w/ ic more than 1


### Functionality - Code style

- recommendation is use `<-` assignment over `->` -- https://github.com/r-earthengine/rgeeExtra/blob/7ea67d064e56e9429971d64440d8424df68f6249/R/ee_gif.R#L187-L192
- in vignettes and examples you seem to embrace the `%>%` syntax for chaining together `{rgee}`/`{rgeeExtra}` processes like this:

```{r}
doi <- ee$ImageCollection("NASA/GPM_L3/IMERG_V06")[[1]] %>%
  ee$Image$Extra_getDOI()
```

rather than:
```{r}
doi <- ee$ImageCollection("NASA/GPM_L3/IMERG_V06")[[1]]$
  Extra_getDOI()
```

is this somehow best practices or you just think easier for R-users? I think the second is certainly easier for users w/ experience in GEE code editor. I also think it's cleaner especially when chaining together many processes.I guess the issue is that it currenlty when using the `$` on ee object you don't get the `rgeeExtra` functions directly. I think this would be a huge improvement if possible.

I'd make it clear somewhere in the documentation that you can do both and when or why you might choose one style over the other.

- can remove leftover code
https://github.com/r-earthengine/rgeeExtra/blob/7ea67d064e56e9429971d64440d8424df68f6249/R/ops-math.R#L108-L110
- In summary methods you rename the new band to "layer" no matter what operation( `summ_r$rename("layer")`)? Would it be safer to explicitly rename the band based on the band_name + Reducer name (i.e B1_max)?
- `raster.R` - why is this file called raster? name seems out of place?
- `ee_get()`- wonder if this name is a bit ambiguous considering all the other `get_*` functions there are in `rgee`,`GEE` and `rgeeExtra`... a couple ideas which make more sense to me filterIndex, filter_index, slice, subset, subset_index



















